Grow with the flow: a latitudinal cline in physiology is associated with more variable precipitation in \textit{Erythranthe cardinalis}

### 2017-09-02 ms/ms.Rnw
added link to github repo
adding mortality effect size analysis

@@

  # Confidence intervals on effect size (mean squared error)
  fn <- function(merMod) {
    x <- anova(merMod)
    ret <- x[c("PopID", "PopID:TempTrt", "PopID:WaterTrt"), "Mean Sq"]
    ret <- c(ret, x["PopID", "Mean Sq"] - x["PopID:TempTrt", "Mean Sq"],
             x["PopID", "Mean Sq"] - x["PopID:WaterTrt", "Mean Sq"])
    ret
  }

@@

  fitMort <- glmer(DiedFromStress ~ PopID * TempTrt * WaterTrt + (1|Family) +
  								 (1|Block), control = glmerControl(optimizer = "bobyqa"),
  								 data = subset(master, !is.na(master$DiedFromStress)), 
  								 family = "binomial")
  # Anova(fitMort, type = 2)

  fn <- function(merMod) {
    x <- Anova(merMod, type = 2)
    ret <- x[c("PopID", "PopID:TempTrt", "PopID:WaterTrt"), "Chisq"]
    ret <- c(ret, x["PopID", "Chisq"] - x["PopID:TempTrt", "Chisq"],
             x["PopID", "Chisq"] - x["PopID:WaterTrt", "Chisq"])
    ret
  }

  system.time(
  ciMerModMort <- confint(fitMort, method = "boot", FUN = fn, nsim = 1e1, seed = 925200709,
                          parallel = "multicore", ncpus = 2L)
)
  # Figure of reaction norms
  d <- subset(master, !is.na(master$DiedFromStress))
  tab <- tapply(d$DiedFromStress, list(d$PopID, d$WaterTrt, d$TempTrt), 
                function(X) 1 - length(which(X)) / length(X))
  pdf(paste(pathFig, "/FigureS_MortRxnNorm.pdf", sep = ""), 4, 4, useDingbats = F)
  par(mai = c(0.5, 1, 0.5, 0.25))
  plot(0, 0, type = "n", xlab = "", ylab = "", axes = F, frame.plot = T, 
       xlim = c(1, 4), ylim = c(0, 1))
  abline(v = 2.5)
  mtext(side = 3, c("Cool", "Hot"), at = c(1.625, 3.375))
  mtext(side = 1, c("Wet", "Dry", "Wet", "Dry"), at = c(1, 2.25, 2.75, 4))
  mtext(side = 1, "Water Treatment", font = 2, cex = 1.25, line = 1.5)
  mtext(side = 3, "Temperature Treatment", font = 2, cex = 1.25, line = 1.5)
  axis(2, lwd = 0, lwd.ticks = 1, las = 1)
  title(ylab = "Survivorship", cex.lab = 1.5)

>>>>

Data and annotated source code to reproduce these analyses and manuscript are available on GitHub (https://github.com/cdmuir/card-cline).

@@

  # Function for bootstrapping effect sizes
  fn <- function(merMod) {
    x <- anova(merMod)
    ret <- x[c("PopID", "PopID:TempTrt", "PopID:WaterTrt"), "Mean Sq"]
    ret <- c(ret, x["PopID", "Mean Sq"] - x["PopID:TempTrt", "Mean Sq"],
             x["PopID", "Mean Sq"] - x["PopID:WaterTrt", "Mean Sq"])
    ret
  }

@@

  fitMort <- glmer(DiedFromStress ~ PopID * TempTrt * WaterTrt + (1|Family) +
  								 (1|Block), control = glmerControl(optimizer = "bobyqa"),
  								 data = subset(master, !is.na(master$DiedFromStress)), 
  								 family = "binomial")
  fit <- Anova(fitMort, type = 2)

  # Function for bootstrapping effect sizes
  fn <- function(merMod) {
    x <- Anova(merMod, type = 2)
    ret <- x[c("PopID", "PopID:TempTrt", "PopID:WaterTrt", 
               "PopID:TempTrt:WaterTrt"), "Chisq"]
    ret <- c(ret, x["PopID", "Chisq"] - x["PopID:TempTrt", "Chisq"],
             x["PopID", "Chisq"] - x["PopID:WaterTrt", "Chisq"],
             x["PopID", "Chisq"] - x["PopID:TempTrt:WaterTrt", "Chisq"])
    ret
  }

  system.time(
  ciMerModMort <- confint(fitMort, method = "boot", FUN = fn, nsim = 1e3, seed = 925200709,
                          parallel = "multicore", ncpus = 4L)
)

  # tmp  <- ciMerModMort
  # ciMerModMort <- tmp
  ciMerModMort <- as.data.frame(ciMerModMort)
  ciMerModMort$estimate <- c(
    fit[c("PopID", "PopID:TempTrt", "PopID:WaterTrt", "PopID:TempTrt:WaterTrt"),
        "Chisq"],
    fit["PopID", "Chisq"] - fit["PopID:TempTrt", "Chisq"],
    fit["PopID", "Chisq"] - fit["PopID:WaterTrt", "Chisq"],
    fit["PopID", "Chisq"] - fit["PopID:TempTrt:WaterTrt", "Chisq"])

      rownames(ciMerModPhoto) <- c("PopID", "PopID:TempTrt", "PopID:WaterTrt", 
                               "PopID - PopID:TempTrt", "PopID - PopID:WaterTrt")
  rm(fit)

  saveRDS(ciMerModMort, paste0(pathDat, "/ciMerModMort.rds"))
  ciMerModMort <- readRDS(paste0(pathDat, "/ciMerModMort.rds"))
  # Figure of reaction norms
  d <- subset(master, !is.na(master$DiedFromStress))
  tab <- tapply(d$DiedFromStress, list(d$PopID, d$WaterTrt, d$TempTrt), 
                function(X) 1 - length(which(X)) / length(X))
  pdf(paste(pathFig, "/FigureS_MortRxnNorm.pdf", sep = ""), 4, 4, useDingbats = F)
  par(mai = c(0.5, 1, 0.5, 0.25))
  plot(0, 0, type = "n", xlab = "", ylab = "", axes = F, frame.plot = T, 
       xlim = c(1, 4), ylim = c(0, 1))
  abline(v = 2.5)
  mtext(side = 3, c("Cool", "Hot"), at = c(1.625, 3.375))
  mtext(side = 1, c("Wet", "Dry", "Wet", "Dry"), at = c(1, 2.25, 2.75, 4))
  mtext(side = 1, "Water Treatment", font = 2, cex = 1.25, line = 1.5)
  mtext(side = 3, "Temperature Treatment", font = 2, cex = 1.25, line = 1.5)
  axis(2, lwd = 0, lwd.ticks = 1, las = 1)
  title(ylab = "Survivorship", cex.lab = 1.5)
